
```{r message=FALSE, warning=FALSE}

#install.packages("tidyverse")
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("plotly")
#install.packages("corrplot")
#install.packages("factoextra")
#install.packages("plyr")
#install.packages("RColorBrewer")
#install.packages("funModeling")
#install.packages("knitr")

library(tidyverse)
library(dplyr)
library(ggplot2)
library(plotly)
library(corrplot)
library(factoextra)
library(plyr)
library(RColorBrewer)
library(funModeling)
library(knitr)
```

loading data
```{r}
set.seed(13232767)
spotify <- read.csv("spotify_songs_test.csv")
```

```{r}
glimpse(spotify)
```

```{r}
colSums(is.na(spotify))
```
Removed NAs
```{r}
spotify <- na.omit(spotify)
```
Removing duplicates
```{r}
spotify <- spotify[!duplicated(spotify$track_id),]
```
Converting key, mode, genre and sub genre to factors 
```{r}
spotify <- spotify %>%
  mutate(playlist_genre = as.factor(spotify$playlist_genre),
         playlist_subgenre = as.factor(spotify$playlist_subgenre),
         mode = as.factor(mode),
         key = as.factor(key))
```
Converting duration_ms to duration in mins (duration_min) 
```{r}
spotify <- spotify %>% mutate(duration_min = duration_ms/60000)
```
new variables that divide popularity into 4 groups for effective cluster analysis 
```{r}
spotify <- spotify %>% 
  mutate(popularity_group = as.numeric(case_when(
    ((track_popularity > 0) & (track_popularity < 20)) ~ "1",
    ((track_popularity >= 20) & (track_popularity < 40))~ "2",
    ((track_popularity >= 40) & (track_popularity < 60)) ~ "3",
    TRUE ~ "4"))
    )
table(spotify$popularity_group)

```
removed track_id, track_album_id and playlist_id from the dataset since it is not useful for our analysis
```{r}
spotify <- spotify %>% select(-c(track_id, track_album_id, playlist_id))
summary(spotify)
```
Correlation Plot
```{r}
df1 <- select(spotify, track_popularity, danceability, energy, loudness, speechiness, acousticness, instrumentalness, liveness, valence, tempo)
corrplot(cor(df1))
```
Histogram: analyzing data distribution of the audio features

```{r}
spotify_hist <- spotify[,-c(1,2,3,4,5,6,7,8,11,13,20,22)]
plot_num(spotify_hist)
```
BOXPLOTS: 

Energy by genre
```{r}
boxplot(energy~playlist_genre, data=spotify,
        main = "Variation of energy between genres",
        xlab = "Energy",
        ylab = "Genre",
        col = "orange",
        border = "brown",
        horizontal = TRUE,
        notch = TRUE
)
```
EDM songs are highest in energy, as expected!

genre by danceability
```{r}
boxplot(danceability~playlist_genre, data=spotify,
        main = "Variation of danceability between genres",
        xlab = "Danceability",
        ylab = "Genre",
        col = "orange",
        border = "brown",
        horizontal = TRUE,
        notch = TRUE
)
```
genre by liveliness
```{r}
boxplot(danceability~playlist_genre, data=spotify,
        main = "Variation of liveness between genres",
        xlab = "Liveness",
        ylab = "Genre",
        col = "orange",
        border = "brown",
        horizontal = TRUE,
        notch = TRUE
)
```
genre by valence
```{r}
boxplot(valence~playlist_genre, data=spotify,
        main = "Variation of valence between genres",
        xlab = "Valence",
        ylab = "Genre",
        col = "orange",
        border = "brown",
        horizontal = TRUE,
        notch = TRUE
)
```
genre by loudness
```{r}
boxplot(loudness~playlist_genre, data=spotify,
        main = "Variation of loudness between genres",
        xlab = "Loudness",
        ylab = "Genre",
        col = "orange",
        border = "brown",
        horizontal = TRUE,
        notch = TRUE
)
```

popularity by accousticness
```{r}
spotify$acousticness.scale <- scale(spotify$acousticness)
spotify %>%
  select(popularity_group, acousticness.scale, playlist_genre) %>%
  group_by(popularity_group)%>%
  filter(!is.na(popularity_group)) %>%
  filter(!is.na(acousticness.scale))%>%
  ggplot(mapping = aes(x = acousticness.scale, y = popularity_group, color = playlist_genre))+
  facet_wrap(~playlist_genre)+
  geom_point()+
  theme_minimal()
```

popularity by valence
```{r}
spotify%>%
  select(popularity_group, valence, playlist_genre) %>%
  group_by(popularity_group)%>%
  filter(!is.na(popularity_group)) %>%
  filter(!is.na(valence))%>%
  ggplot(mapping = aes(x = popularity_group, y = valence, color = playlist_genre, fill = playlist_genre))+
  geom_bar(stat = 'identity')+
  coord_polar()+
  facet_wrap(~playlist_genre)+
  theme_minimal()
```
energy distribution
```{r}
spotify$cut_energy <- cut(spotify$energy, breaks = 10)
spotify %>%
  ggplot( aes(x=cut_energy ))+
  geom_bar(width=0.2) +
  coord_flip() +
  scale_x_discrete(name="Energy")  
```
speechiness distribution
```{r}
spotify$cut_spe <- cut(spotify$speechiness, breaks = 10)
spotify %>%
  ggplot( aes(x=cut_spe ))+
  geom_bar(width=0.2) +
  coord_flip() +
  scale_x_discrete(name="Spechiness")  
```
tempo and liveness distribution across genre
```{r}
spotify$liveness.scale <- scale(spotify$liveness)
spotify$tempo.scale <- scale(spotify$tempo)
spotify %>%
  select(tempo.scale, liveness.scale, playlist_genre) %>%
  group_by(playlist_genre)%>%
  filter(!is.na(tempo.scale)) %>%
  filter(!is.na(liveness.scale))%>%
  ggplot(mapping = aes(x = tempo.scale, y = liveness.scale, color = playlist_genre, fill = playlist_genre))+
  geom_bar(stat = 'identity')+
  coord_polar()+
  theme_minimal()

```

3D Scatter Plot!!!
popularity vs danceability vs energy
```{r}
spotify_arrange <- spotify[order(-spotify$track_popularity),]

 

plot_ly(head(spotify_arrange, n=500), x = ~track_popularity, y = ~danceability, z = ~energy, 
        color = ~playlist_genre, colors = c('magenta', 'green', 'red', 'cyan', 'orange', 'yellow'),size = I(100)) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'track_popularity'),
                      yaxis = list(title = 'danceability'),
                      zaxis = list(title = 'energy')),
         title = "3D Scatter plot: Track_popularity vs Danceability vs Energy",
         showlegend = FALSE)
```

popularity vs genre vs subgenre
```{r}
spotify_arrange <- spotify[order(-spotify$track_popularity),]

 

plot_ly(head(spotify_arrange, n=500), x = ~track_popularity, y = ~playlist_genre, z = ~playlist_subgenre, 
        color = ~playlist_genre, colors = c('magenta', 'green', 'red', 'cyan', 'orange', 'yellow'),size = I(100)) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'track_popularity'),
                      yaxis = list(title = 'playlist_genre'),
                      zaxis = list(title = 'playlist_subgenre')),
         title = "3D Scatter plot: Track_popularity vs Playlist genre vs Playlisty subgenre",
         showlegend = FALSE)
```

top 10 artists and most famous tracks
```{r}
spotify %>%
  select(track_name, track_artist, track_album_name, playlist_genre, track_popularity)%>%
  group_by(track_artist)%>%
  filter(!is.na(track_name))%>%
  filter(!is.na(track_artist))%>%
  filter(!is.na(track_album_name))%>%
  arrange(desc(track_popularity))%>%
  head(n = 10)%>%
  ggplot(mapping = aes(x = track_name, y =  track_artist, color = track_artist, fill = track_artist, size = track_popularity ))+
  geom_point()+
  coord_polar()+
  facet_wrap(~playlist_genre)+
  theme_minimal()+
  labs(x = 'track_name', y = 'track_artist', title = 'Top ten artists of spotify')+
  theme(plot.title = element_text(hjust=0.5),legend.position ='bottom')
```

CLUSTERING!
Variables used: danceability, energy, loudness, speechiness, acousticness, instrumentalness, liveness, valence, tempo, duration_min

Checking the updated dataset for EDA:
```{r}
str(spotify)
```
Scaling the numeric variables required for cluster analysis:
```{r}
spotify_scaled <- scale(spotify[,-c(1,2,3,4,5,6,7,8,11,13,20,22,23,24,25,26,27)])
summary(spotify_scaled)
```

Determining the optimal number of clusters
```{r}
wss <- function(data, maxCluster = 9) {
  SSw <- (nrow(data) - 1) * sum(apply(data, 2, var))
  SSw <- vector()
  for (i in 2:maxCluster) {
    SSw[i] <- sum(kmeans(data, centers = i)$withinss)
  }
  plot(1:maxCluster, SSw, type = "o", xlab = "Number of Clusters", ylab = "Within groups sum of squares", pch=19)
}

wss(spotify_scaled)
```
k = 7, this seems optimal

K-means Clustering:
```{r}
spotify_kmeans <- kmeans(spotify_scaled, centers = 7)
spotify_kmeans$size

spotify_kmeans$centers

spotify$cluster <- spotify_kmeans$cluster
tail(spotify)
```
Plotting using 'factoextra':

```{r}
fviz_cluster(spotify_kmeans, data=spotify_scaled)
```

goodness of fit tests, 3 values ot check:

1.Within Sum of Squares tot.withinss : signifies the 'length' from each observation to its centroid in each cluster

```{r}
spotify_kmeans$tot.withinss
```

2. Total Sum of Squares totss : signifies the 'length' from each observation to global sample mean

```{r}
spotify_kmeans$totss
```

3. Between Sum of Squares betweenss : signifies the 'length' from each centroid from each cluster to the global sample mean

```{r}
spotify_kmeans$betweenss
```

Another 'goodness' measure can be signifies with a value of betweenss/totss closer the value to 1 or 100%, the better): `betweenss`/`tot.withinss`

```{r}
((spotify_kmeans$betweenss)/(spotify_kmeans$totss))*100
```

Good cluster has high similarity characteristics in 1 cluster (low WSS) and maximum difference in characteristics between clusters (high BSS). In addition, it can be marked with a BSS / totss ratio that is close to 1 (100%).

From the unsupervised learning analysis above, we can summarize that K-means clustering can be done using this dataset since we have got a reasonable high value for BSS / totss ratio, 40.78%.

Repeating the same exercise multiple times by adjusting with multiple combinations of variables, we are getting the best fit and optimized model by excluding loudness, tempo and duration_min.

Finding what kind of song characterises each clusters in the optimized model:
```{r,warning=FALSE}
spotify %>% 
  group_by(cluster) %>% 
  summarise_all(mean) %>% 
  select(cluster, acousticness, danceability, energy, instrumentalness, speechiness, valence, liveness)
```
 
Characteristics:
- Cluster 1: Highest danceability, Highest valence
- Cluster 2: Highest energy, Lowest acousticness
- Cluster 3: Lowest instrumentalness, Highest speechiness
- Cluster 4: Lowest speechiness, Lowest liveness
- Cluster 5: Lowest acousticness, Highest liveness  
- Cluster 6: Highest instrumentalness
- Cluster 7: Highest acousticness, Lowest energy, Lowest valence

SONG RECOMMENDATIONS

Now, let's check the which cluster is my favorite song. My favourite track from the list is Memories by Maroon 5.

```{r}
spotify %>% 
  filter(track_name == "Memories - Dillon Francis Remix", track_artist == "Maroon 5")
```

my favorite song is in cluster 3.

Now I want to try a new genre, r&b. So let's check the best songs which I should try according to my taste in the r&b genre given that my favourite song is Memories by Maroon 5.

```{r}
spotify %>% 
  filter(cluster == 3, playlist_genre == "r&b") %>% 
  sample_n(5)
```
